---
title: "Copula"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(readr)
library(dynlm)
library(normtest)
library(car) 
library(lmtest)
library(REndo)
library(MLmetrics)
library(gcmr)
options(scipen=4)

library(plm) # plm
library(gplots) # plotmeans
library(psych) # describeBy
```

```{r}
###--Testing model for problems
testing.model <- function(model) {
  cat("--==Testing model==--", fill = T)
  cat(fill = T)
  cat("R-squared:", summary(model)$r.squared, fill = T)
  cat("Adj. R-squared:", summary(model)$adj.r.squared, fill = T)
  cat("AIC: ", AIC(model), fill = T)
  cat(fill = T)
  cat("Normality of disturbances: ", jb.norm.test(model$residuals)$p.value, fill = T)
  cat("Multicollinearity: ", max(vif(model)), fill = T)
  cat("Heteroskedasticity: ", bptest(model)$p.value, fill = T)
  cat("Autocorrelation of 1 order: ", bgtest(model)$p.value, fill = T)
  cat("RESET-test: ", reset(model)$p.value, fill = T)
}
```

```{r}
###--Choosing the panel model (RE-FE-pooled)
choose_panel <- function(){
  # F-test (Pooled VS FE-model)
  if (pFtest(formulaP, dataP, type="individual")$p.value>=0.05){
    # Breusch-Pagan test (Pooled VS RE-model)
    if (plmtest(plm(formulaP, dataP, model="pooling"), 
                effect="twoways", type="bp")$p.value>=0.05){
      return("Choose pooled model")
    }else{
      return("Choose RE-model")}
  }else{
    # Hausman-test (RE-model VS FE-model)
    if (phtest(plm(formulaP, dataP, model="within"), 
               plm(formulaP, dataP, model="random", 
                   random.method = "walhus"))$p.value>=0.05){
      return("Choose RE-model")
    }else{
      return("Choose FE-model")
    }
  }
}
```

# Russia

##--Importing data

```{r}
dataRF <- read_csv("dataRF.csv", col_types = cols(Date = col_character()))
dataRF$perSIDinNGDP <- dataRF$SID*100/(dataRF$NGDP*4)
dataRF$LUNEMPL <- c(5.9, dataRF$UNEMPL[-45])
dataRF$LNPLp <- c(1.76780823887159, dataRF$NPLp[-45])
```

##--OLS model

```{r}
formula <- NPLp~
            #UNEMPL + 
            #I(UNEMPL^2)+
            #bvCrisis+
            perSIDinNGDP+
            indexDemocracy+
            indexInstitute+
            indexPayment+
            LNPLp + 
            bvQ1 + 
            bvQ2 +
            bvQ3
mdlRF <- lm(data=dataRF, formula=formula)
summary(mdlRF)
testing.model(mdlRF)
cat("MAPE: ", MAPE(y_true=dataRF$NPLp, y_pred=mdlRF$fitted.values),fill = T)
```


##--Copula regression

```{r}
# Adjust randomness
set.seed(1234)
# Upgrade initial OLS formula to copula
formulaCPL <- formula(paste(formula[2], formula[1], formula[3],
                   " | continuous(perSIDinNGDP, indexPayment, indexInstitute)"))
cplRF <- copulaCorrection(formula =  formulaCPL, num.boots = 10000, data=dataRF)
```
```{r}
summary(cplRF)
```

```{r}
jb.norm.test(cplRF$residuals)
acf(cplRF$residuals)
bptest(cplRF)
bgtest(cplRF, order = 1)
bptest(cplRF)
reset(cplRF)
cat("MAPE: ", MAPE(y_true=dataRF$NPLp, y_pred=cplRF$fitted.values),fill = T)
```


##--Comparison of simulation results

```{r}
tab <- matrix(c(c(mdlRF$coefficients, rep(NA,3)), cplRF$coefficients), ncol=2)
colnames(tab) <- c("OLS", "Copula")
rownames(tab) <- cplRF$names.main.coefs
tab
```


# Kazakhstan

##--Importing data

```{r}
dataKZ <- read_csv("dataKZ.csv", col_types = cols(Date = col_character()))
dataKZ$perSIDinNGDP <- dataKZ$SID*100/dataKZ$NGDP
dataKZ$LUNEMPL <- c(5.5, dataKZ$Unempl[-36])
dataKZ$LNPLp <- c(14.6071858314849, dataKZ$NPLp[-36])
```


##--OLS model

```{r}
formula <- NPLp ~ 
  #LUNEMPL + 
  perSIDinNGDP + 
  indexDemocracy + 
  indexInstitute + 
  indexPayment +
  #COFS+
  Slr+
  #AGRO+
  #I(GREV/NGDP)+
  #taxProfit+
  #CH+
  #I(SED/NGDP)+
  indexBureau+
  #RealInc+
  #CM+
  #DR+
  LNPLp + 
  bvQ1 + 
  bvQ2 +
  bvQ3
mdlKZ <- lm(data=dataKZ, formula = formula)
summary(mdlKZ)
testing.model(mdlKZ)
```

##--Copula regression

```{r}
set.seed(1234)

formulaCPL <- formula(paste(formula[2], formula[1], formula[3],
                   " | continuous(perSIDinNGDP, indexPayment, indexInstitute)"))
cplKZ <- copulaCorrection(formula =  formulaCPL, num.boots = 10000, data=dataKZ)
summary(cplKZ)
```

```{r}
jb.norm.test(cplKZ$residuals)
acf(cplKZ$residuals)
bptest(cplKZ)
bgtest(cplKZ, order = 2)
bptest(cplKZ)
reset(cplKZ)
```

##--Comparison of simulation results

```{r}
tab <- matrix(c(c(mdlKZ$coefficients, rep(NA,3)), cplKZ$coefficients), ncol=2)
colnames(tab) <- c("OLS", "Copula")
rownames(tab) <- cplKZ$names.main.coefs
tab
```


# Belarus

##--Importing data

```{r}
dataRB <- read_csv("C://R//Data//PaperGit//Paper//dataRB.csv", col_types = cols(Date = col_character()))
dataRB$perSIDinNGDP <- dataRB$SID*100/(dataRB$NGDP)
dataRB$LNPLp <- c(1.09, dataRB$NPLp[-45])
dataRB$lgGREV <- log(dataRB$GREV)
dataRB$lgGEXP <- log(dataRB$GEXP)
```


##--OLS model

```{r}
colnames(dataRB)
```


```{r}
formula <- NPLp ~ 
  #iloUnempl + 
  #I(iloUnempl^2) + 
  #I(perSIDinNGDP^2) + 
  perSIDinNGDP+
  #indexInstitute + 
  indexBureau + 
  #indexPayment+ 
  #indexDemocracy + 
  lgGREV +
  lgGEXP+
  #Tax+ 
  LNPLp+
  bvQ1 + 
  bvQ2 + 
  bvQ3 
mdlRB <- lm(data=dataRB, formula = formula)
summary(mdlRB)
testing.model(mdlRB)
```



##--Copula regression

```{r include=FALSE}

# Adjust randomness
set.seed(1234)
formulaCPL <- formula(paste(formula[2], formula[1], formula[3],
                   " | continuous(perSIDinNGDP, indexBureau)"))

cplRB <- copulaCorrection(formula =  formulaCPL, num.boots = 10000, data=dataRB)
```

```{r}
summary(cplRB)
```

```{r}
jb.norm.test(cplRB$residuals)
acf(cplRB$residuals)
bptest(cplRB)
bgtest(cplRB, order = 2)
reset(cplRB)
```

##--Comparison of simulation results

```{r}
tab <- matrix(c(c(mdlRB$coefficients, rep(NA,2)), cplRB$coefficients), ncol=2)
colnames(tab) <- c("OLS", "Copula")
rownames(tab) <- cplRB$names.main.coefs
tab
```


# Kyrgyzstan

##--Importing data

```{r}
dataKG <- read_csv("C://R//Data//PaperGit//Paper//dataKG.csv", col_types = cols(Date = col_character()))
dataKG$perSIDinNGDP <- dataKG$SID*100/dataKG$NGDP
```

##--OLS model

```{r}
colnames(dataKG)
```


```{r}
formula <- NPLp ~ 
  perSIDinNGDP + 
  #I(GREV*100/NGDP) + 
  #I(GEXP*100/NGDP) +
  #WSGDP +
  #Tax+
  #DCPS+
  #AGRO+
  #DR+
  #CM+
  #CH+
  #vTax+
  I(Unempl^2) + 
  indexBureau  + 
  indexDemocracy + 
  indexInstitute + 
  indexPayment + 
  bvQ1 + 
  bvQ2 +
  bvQ3
mdlKG <- lm(data=dataKG, formula = formula)
summary(mdlKG)
testing.model(mdlKG)
```

##--Copula regression

```{r include=FALSE}

# Adjust randomness
set.seed(1234)
formulaCPL <- formula(paste(formula[2], formula[1], formula[3],
                   " | continuous(perSIDinNGDP, indexPayment, indexInstitute)"))
cplKG <- copulaCorrection(formula =  formulaCPL, num.boots = 10000, data=dataKG)
```

```{r}
summary(cplKG)
```

##--Comparison of simulation results

```{r}
tab <- matrix(c(c(mdlKG$coefficients, rep(NA,3)), cplKG$coefficients), ncol=2)
colnames(tab) <- c("OLS", "Copula")
rownames(tab) <- cplKG$names.main.coefs
tab
```

# Armenia

## --Importing data

```{r}
dataAR <- read_csv("C://R//Data//PaperGit//Paper//dataAR.csv", col_types = cols(Date = col_character()))
dataAR$perSIDinNGDP <- dataAR$SID*100/dataAR$NGDP
dataAR$LNPLp <- c(NA, dataAR$NPLp[-44])
dataAR$L2NPLp <- c(NA, dataAR$LNPLp[-43])
dataAR$LUnempl <- c(NA, dataAR$Unempl[-44])

dataAR <- dataAR[3:44,]
```

##--OLS model

```{r}
colnames(dataAR)
```


```{r}
formula <- NPLp ~ 
  perSIDinNGDP + 
  #I(GREV*100/NGDP) + 
  #I(GEXP*100/NGDP) +
  #WSGDP +
  #Tax+
  #DCPS+
  #AGRO+
  #DR+
  #CM+
  #CH+
  #vTax+
  Unempl +
  indexBureau  + 
  indexDemocracy + 
  indexInstitute +
  LNPLp+
  L2NPLp+
  bvQ1 + 
  bvQ2 +
  bvQ3
mdlAR <- lm(data=dataAR, formula = formula)
summary(mdlAR)
testing.model(mdlAR)
```


##--Copula regression

```{r include=FALSE}

# Adjust randomness
set.seed(1234)
formulaCPL <- formula(paste(formula[2], formula[1], formula[3],
                   " | continuous(perSIDinNGDP, indexInstitute)"))
cplAR <- copulaCorrection(formula =  formulaCPL, num.boots = 10000, data=dataAR)
```

```{r}
summary(cplAR)
```


##--Comparison of simulation results

```{r}
tab <- matrix(c(c(mdlAR$coefficients, rep(NA,2)), cplAR$coefficients), ncol=2)
colnames(tab) <- c("OLS", "Copula")
rownames(tab) <- cplAR$names.main.coefs
tab
```



# Panel OLS

##--Importing and formatting the data

```{r}
columns <- c("Date", "Country", "NGDP", "WSGDP", "Tax", "DR", "AGRO", "GREV", 
             "GEXP", "indexInstitute", "indexBureau", "indexPayment", 
             "indexDemocracy", "Unempl", "bvQ1", "bvQ2", "bvQ3", "M2", "M0",
             "CH", "NPLa", "NPLp", "SID", "CM", "perSIDinNGDP", "LNPLp", 
             "CountryID")

# Russia
dataRF <- read_csv("C://R//Data//PaperGit//Paper//dataRF.csv", col_types = cols(Date = col_character()))
dataRF$perSIDinNGDP <- dataRF$SID*100/dataRF$NGDP
dataRF$LUNEMPL <- c(5.9, dataRF$UNEMPL[-48])
dataRF$LNPLp <- c(1.76780823887159, dataRF$NPLp[-48])
dataRF <- dataRF[9:48,]
colnames(dataRF)[19] <- "Unempl"
dataRF$Country <- c(rep("Russia", 40))
dataRF$CountryID <- c(rep(0, 40))
dataRF <- dataRF[columns]

# Kazakhstan
dataKZ <- read_csv("C://R//Data//PaperGit//Paper//dataKZ.csv", col_types = cols(Date = col_character()))
dataKZ$perSIDinNGDP <- dataKZ$SID*100/dataKZ$NGDP
dataKZ$LUNEMPL <- c(5.5, dataKZ$Unempl[-40])
dataKZ$LNPLp <- c(14.6071858314849, dataKZ$NPLp[-40])
dataKZ$Country <- c(rep("Kazakhstan", 40))
dataKZ$CountryID <- c(rep(1, 40))
dataKZ <- dataKZ[columns]

# Belarus
dataRB <- read_csv("C://R//Data//PaperGit//Paper//dataRB.csv", col_types = cols(Date = col_character()))
dataRB$perSIDinNGDP <- dataRB$SID*100/dataRB$NGDP
dataRB$LNPLp <- c(1.09, dataRB$NPLp[-45])
dataRB <- dataRB[6:45,]
colnames(dataRB)[17] <- "Unempl"
dataRB$Country <- c(rep("Belarus", 40))
dataRB$CountryID <- c(rep(2, 40))
dataRB <- dataRB[columns]

# Kyrgyzstan
dataKG <- read_csv("C://R//Data//PaperGit//Paper//dataKG.csv", col_types = cols(Date = col_character()))
dataKG$perSIDinNGDP <- dataKG$SID*100/dataKG$NGDP
dataKG$LNPLp <- c(NA, dataKG$NPLp[-49])
dataKG <- dataKG[10:49,]
dataKG$Country <- c(rep("Kyrgyzstan", 40))
dataKG$CountryID <- c(rep(3, 40))
dataKG <- dataKG[columns]
rm(columns)

# Applying to the panel
dataPanel <- rbind(dataRF, dataKZ, dataRB, dataKG)
dataPanel <- pdata.frame(dataPanel, index=c("Country", "Date"))
rm(dataRF, dataKZ, dataRB, dataKG)
```

```{r}
pdim(dataPanel)
summary(dataPanel)
```

##--Entering the formula

```{r}
frm <- NPLp~Unempl + I(Unempl^2)+perSIDinNGDP+indexDemocracy+indexInstitute+indexPayment+indexBureau+LNPLp + bvQ1 + bvQ2 +bvQ3
```

##--Choosing the panel model (RE-FE-pooled) 

```{r}
dataP= dataPanel
formulaP= frm
choose_panel()
rm(dataP, formulaP)
```

##--The pooled model estimation

```{r pooled model}
mdlPooled <- plm(frm, data=dataPanel, model="pooling")
summary(mdlPooled)
testing.model(mdlPooled)
# the Wooldridge's test for unobserved effects
pwtest(mdlPooled, effect = c("individual"))
```

##--FE-model

```{r FEmodel}
mdlFE <- plm(frm, data=dataPanel, model="within")

# FE-model summary 
summary(mdlFE)

# The Wald test for overall significance
pwaldtest(mdlFE)

acf(mdlFE$residuals)
jb.norm.test(mdlFE$residuals)

# extract the fixed effects
summary(fixef(mdlFE))
```


##--RE-model

```{r REmodel}
mdlRE <- plm(frm, data=dataPanel, model="random", random.method = "walhus")

# RE-model summary 
summary(mdlRE)

# The Wald test for overall significance
pwaldtest(mdlRE)
```

# Panel copulas

```{r}
frm <- NPLp~Unempl + I(Unempl^2)+perSIDinNGDP+indexDemocracy+indexInstitute+indexPayment+indexBureau+LNPLp + bvQ1 + bvQ2 +bvQ3 
```

```{r}
mdlCopulaPanel <- gcmr(frm, dataPanel, marginal = gaussian.marg,
                       cormat = cluster.cormat(CountryID, "ar1"),
     options=gcmr.options(seed=123, nrep=10000))
```

```{r}
summary(mdlCopulaPanel)
```

